### 2020-10-14 generated/tables/table_time.tex
grid7

\begin{table}

\caption{\label{tab:table_time}Average runtime in hours, for evaluating one grid and one optimization run of MBO.}
\centering
\begin{tabular}[t]{lrrrrrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4.5}{c}{linear} & \multicolumn{4.5}{c}{paper} & \multicolumn{4.5}{c}{paper2} & \multicolumn{4.5}{c}{sigmoid} \\
\cmidrule(l{3pt}r{3pt}){2-5.5} \cmidrule(l{3pt}r{3pt}){6.5-10} \cmidrule(l{3pt}r{3pt}){11-14.5} \cmidrule(l{3pt}r{3pt}){15.5-19}
 & 500 & 1000 & 2000 & 500 & 1000 & 2000 & 500 & 1000 & 2000 & 500 & 1000 & 2000\\
\midrule
Grid & 55.1 & 60.9 & 60.5 & 62.7 & 63.2 & 66.6 & 59.7 & 64.4 & 67.1 & 54.6 & 52.8 & 61.0\\
Grid 7 & 3.4 & 3.8 & 3.7 & 3.8 & 3.9 & 4.1 & 3.8 & 3.9 & 4.1 & 3.4 &  & 3.7\\
MBO & 3.6 & 3.4 & 3.6 & 3.4 & 4.0 & 4.1 & 3.7 & 3.3 & 3.2 & 3.3 & 2.6 & 3.1\\
\bottomrule
\end{tabular}
\end{table}