### 2020-11-02 generated/tables/table_time.tex
final results

\begin{table}

\caption{\label{tab:table_time}Average runtime in hours, for evaluating one grid and one optimization run of MBO.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{paper} & \multicolumn{3}{c}{paper2} & \multicolumn{3}{c}{sigmoid} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
 & 500 & 1000 & 2000 & 500 & 1000 & 2000 & 500 & 1000 & 2000 & 500 & 1000 & 2000 & \emph{evals}\\
\midrule
Grid & 51.2 & 55.3 & 56.3 & 56.7 & 58.7 & 63.0 & 55.1 & 59.2 & 61.6 & 50.7 & 51.3 & 56.8 & 1350\\
Grid Small & 3.8 & 4.1 & 4.1 & 4.1 & 4.2 & 4.6 & 4.1 & 4.3 & 4.4 & 3.7 & 3.8 & 4.1 & 126\\
MBO & 3.5 & 3.3 & 3.7 & 3.3 & 3.6 & 3.7 & 3.5 & 3.3 & 3.2 & 3.3 & 2.9 & 3.6 & 116\\
\bottomrule
\end{tabular}}
\end{table}